A Recipe for Scavenging - the natural history of a behaviour

### 2016-04-08 natural history of scavenging.tex
bipedalism and swimming locomotion

%KH: need paragraph for how do we know something is (was) a scavenger
\section*{Encounter Rate}
All foraging processes depend on the encounter rate between the consumer and resources. 
In the simplest case, this rate can be thought of in terms of a gas diffusion model where the movement of two agents depends only on their relative speed.
As carcasses are stationary, the relative speed between a scavenger and carrion is only dependent on the movement of the scavenger. 
Vertebrates do more than simply bump into resources like gas molecules though, because they can actively detect them through their sensory abilities. 
As such, scavenging potential is strongly affected by search rates, which are determined by both species' physiology and the dimensionality of the environment \citep{pawar2012dimensionality}.

\subsection*{Locomotion}
Scavenging depends more on the ability to efficiently move over large areas than does predation because of the inherent unpredictability of carrion. 
This generally requires an efficient transfer of metabolic energy into movement which relies on both physiology (i.e. metabolism) and the medium of the environment in which the animal is moving (i.e. aerial, aquatic or terrestrial).
Perhaps the most efficient form of locomotion in vertebrates is, paradoxically, found in flying species. 
Despite the energetic costs of flight, the only known vertebrate obligate scavengers are the old and the new world vultures. 
And, although powered flight is energetically expensive, species like vultures can exploit air currents using their large wingspans which allows them to soar at a cost of only twice their metabolic rate \citep{hedenstrom1993migration,spivey2014analysing}. %TG: I didn't get the meaning of this sentence, is soaring doubling the metabolic rate? or is it a word misplacement ("while only doubling their metabolic rate")? and if so, how does that make the point?
By depending on thermal air flows these species can forage over vast ranges \citep{spiegel2013factors}. 
An analagous mode of locomotion is also exploited by sea birds, which use strong ocean winds to search large areas of the oceans \citep{norberg2012vertebrate,thaxter2012seabird}. 
While many species of seabird are likely primarily predators, it seems that albatrosses, who can range many hundreds of kilometres, take a substantial amount of carrion in their diet \citep{croxall1994dead}. 
This is typically in the form of squid, which float on the surface, allowing the birds to readily pluck their remains out of the water \citep{croxall1994dead}.

@@

Similar to aerial species, aquatic scavengers have a locomotory benefit because water is a medium that is conducive to low-cost movement \citep{tucker1975energetic}. 
In fact, the total cost of movement (again in J kg\textsuperscript{-1} m\textsuperscript{-1}) in fish is lower than either running or flying where it scales according to 2.15 $\times$ body mass (kg) \textsuperscript{-0.25} \citep{williams1999evolution} with only \textit{soaring} flight likely to surpass it. 
%KH(I assume the reference is for powered flight, again this needs to be more precise).  
This has led some researchers to argue for the likelihood of an obligate scavenging fish \citep{ruxton2004energetic,ruxton2005searching}. 
Sharks perhaps best resemble the large soaring fliers as they depend on large pectoral fins in order to maintain lift as they swim. 
Many shark species show very large ranges (e.g. great white sharks (\textit{Carcharodon carcharias})) and have the lowest cost of swimming. %TG: I guess the following is a comment: (I know the Greenland shark is the slowest swimmer). 
Facultative scavenging is seen in many marine groups including species of extant sharks such as great white sharks \citep[known to feed on whale carcasses;][]{fallows2013white}, Greenland sharks \citep[feeding on seals;][]{watanabe2012slowest}, and sixgill sharks \citep{anderson2016impact}.
There is evidence too of scavening in extinct species, where shark teeth have been found in the remains of dinosaurs, mosasaurs and mysticete whales from the Pliocene \citep[5.3 to 3.6 Mya;]{schwimmer1997scavenging,ehret2009caught}.

@@

%% AK:  Need to be careful here because this whole idea of sprawling posture being worse off than erect posture is not that well supported in the literature. 
%% Great Transformations in Vertebrate Evolution By Kenneth P. Dial, Neil Shubin, Elizabeth L. Brainerd talks about it at length 
%% So I think this should stay pitched as an endo/ ectothermy thing rather than posture 
Terrestrial environments are the most energetically costly in which to move, because an animal must trade off the cost of supporting its body off the ground while propelling itself forward. 
The idea that the evolutionary transition in posture from the sprawling gait of reptiles to the erect posture of mammals conferred a huge advantage to the latter is often mentioned.
The purported advantages include benefits in terms of speed, efficiency, muscle effort and manoeuvrability \citep{sullivan2015posture}. 
Yet, as \cite{sullivan2015posture} states, most of the hypotheses in favour of this idea remain to be tested in the context of archosaur evolution. 
Metabolic rate however, unquestionably impacts terrestrial species whereby ectotherms such as many modern reptiles, cannot move for sustained periods \citep{bennett1979endothermy}. 
% Species which do not support an upright position such as modern reptiles limit their ability to move for sustained periods \citep{bennett1979endothermy}. 
This is exacerbated by their sprawling gait which results in the phenomenon known as Carrier's constraint such that the animal can't move and breathe at the same time because the lateral movements impedes its lungs \citep{carrier1987evolution}. 
% These constraints are likely to mean that many ectotermic terrestrial species, which cannot maintain an upright posture would depend on a relatively low level of scavenging (I'm not sure this is true though). 
This would also be true of extinct species. 
It is during the Permian (298.9 - 252.17 Mya) that we have the earliest evidence of vertebrate scavenging where a temnospondyl amphibian fed on the carcass of \textit{Varanops}, a predatory synapsid of the time \citep{reisz2006articulated}. 
%TG: I guess this is a comment too: Snakes are a bit different are more effecent at moving I think - lots of scavanging. 
Both these groups are more likely to depend on low energetic costs in order to compensate for low encounter rates due to their inability to move. 
It is with the evolution of endothermy in the therapsid-mammal lineage \citep{clarke2010temperature} that terrestrial vertebrates would have gained the ability to range more widely, a vital component in seeking out carrion.

Modern endothermic mammals can sustain longer periods of energetically expensive activity \citep{bennett1979endothermy} resulting in larger foraging ranges. 
To quantify this effect with a simple example we can turn to some allometric relationships relating sustainable travelling speed to body mass \citep{ruxton2004obligate}.
If we insert these into a foraging radius model \citep{Enstipp2006Energetics} for a 12 hour foraging day it shows that while a 10 kg reptile can range 6.5 km, an equally sized mammal can range nearly 33 km . 
For a foraging scavenger, this ability translates into a greater area searched for food.

@@

%TG: missing some transition here. Maybe something short like "Further back in time" or "Previous to that"...
The evolution of bipedal movement freed up the forelimbs for other purposes e.g. wings, tool-use etc. 
In terms of its cost \cite{alexander2004bipedal} shows that, in the case of humans, we are more economical than predicted while walking and less so while running.

%  KH Something more a bipedal movement.
Unsurprisingly, given their enduring appeal, the prevalence of scavenging has been explored in the carnivorous, theropod dinosaurs.
They were the dominant terrestrial carnivores for most of the Mesozoic Era (252.17 - 66 Mya) and ranged from the chicken-sized to the whale-sized, all of which were bipedal.
They are quite alien to anything we know today which restricts our ability to understand their ecology far more so than extinct mammals \citep{weishampel2004dinosauria}.
Of relevance, are the questions that still persist about their metabolism, with the latest evidence suggesting they were mesothermic i.e. intermediate to ecto- and endotherms \citep{grady2014evidence}. 
We do know that they walked with the erect gait of mammals or birds rather than the sprawling gait of lizards and that they were most likely facultative scavengers \citep{weishampel2004dinosauria,depalma2013physical}. %TG: and cite "Kane et al in press" no? I think these people did some kind of work on that matter.
Taken together, this implies dinosaurs had a foraging range that fell in between that of modern terrestrial mammals and reptiles.

@@

\subsection*{Detection}
It would be pointless to have incredible ranging abilities and not have the sensory architecture to benefit from it.
%If we came at this from a position of complete ignorance we would predict scavengers to have well-developed senses and indeed, this is what we find. %TG: I don't really like this sentence. How about:
As predicted by the necessity of an increased encounter rate, scavengers have well-developed detection senses.
A simplification of terrestrial, vertebrate scavengers in sensory terms is one of them existing in a two-dimensional plane while foraging for carrion directly.
They can detect carcasses at a range that is defined by the radius of their sensory organs. %, usually the visual and olfactory senses.
As a consequence, they have a much more restricted view of the landscape than do aerial foragers.
Hyenas make up for this in their ability to smell a rotting carcass 4 km away and to hear the vocalisations of conspecifics at a distance of 10 km \citep{mills1989comparative}. 
% Using the approach of \cite{spiegel2013factors} we estimate a spotted hyena %TG: do we need a (\textit{Crocuta crocuta}) here?
%could resolve a 2 metre target at 1 km distance. 
If we compare this to the energetics approach by \cite{ruxton2004obligate}, they calculated a terrestrial scavenger needs to be able to detect carrion at 500 meter in order to survive, which is clearly within the ability of hyenas.  
% While considering prehistoric habitats \cite{ruxton2004obligate} calculated that ``a 1 tonne mammal or reptile, in an ecosystem yielding carrion at densities similar to the current Serengeti, could have met its energy requirements if it could detect carrion over a distance of the order of 400–500 m".
Moreover, the senses of many extant (and in all probability extinct) carnivores meet this required distance, making scavenging feasible for terrestrial species \citep{farlow1994speculations,mech2010wolves}.

@@

Having a panoramic view also means being able to gather a wealth of information from other foragers, be they conspecifics or otherwise \citep{jackson2008effect}.
Again, returning to vultures, the genus \textit{Gyps} consists of highly social and colonially nesting species \citep{fernandez2015density}.
These behaviours allow them forage far more efficiently because one bird can scrounge information on the location of food from another successful forager \citep{KaneVul}.

@@

\section*{Handling Time}
Since carrion is not dispatched directly, often the most easily accessible and choicest components of the carcass will be missing or, if present, will be fought over.
Being able to extract nutrients from remnants gives a scavenger a great advantage.
%Thus, the bone crushing ability of hyenas and others reveals another useful scavenger trait. %TG: how about the following for a smoother transition:
Thus, some phenotypic adaptations such as the ability to crush bone (as seen in hyenas) reveals another useful scavenger trait.
Osteophagy is known across a range of terrestrial carnivores and given that some fat-rich mammalian bones have an energy density (6.7 kJ/g) comparable with that of muscle tissue, it makes skeletal remains an enticing resource \citep{brown1989study}.
This ability reached its zenith among hyenas with the evolution of the 110 kg \textit{Pachycrocuta brevirostris} during the Pliocene \citep[3.6 - 2.58 Mya; ][]{palmqvist2011giant}.
Indeed, their extinction has been blamed on the decline of sabretooth cats (Machairodontinae), the unique skull morphology of the latter meant they would leave a large amount of food on a carcass for would-be scavengers \citep{palmqvist2011giant}. 
Earlier in the evolution of mammals, the bone-crushing dogs that evolved during the Oligocene (Borophaginae; 33.9 - 23.03 Mya) have also been compared to hyenas in terms of their feeding ecology \citep{van2003chapter,martin2016pursuit}.

@@

Interestingly, such comparisons have given insight into the feeding ecology of early hominins who, for instance, had the ability to craft tools for breaking open bones \citep{ARCM:ARCM12084}.
The question of where our ancestors placed on the hunter-scavenger axis during the Plio-Pleistocene has been a matter of debate for years \citep{dominguez2002hunting}.
A recent study investigating potential scavenging opportunities for hominins in Kenya found that, even when discounting bone material, there is a substantial amount of scavengeable %TG: is that an adjective?
meat left on predated remains; sufficient to sustain the requirements of an adult male \textit{Homo erectus} \citep{pobiner2015new}.
In some historical hominin-inhabited areas there were a greater number of felids than hyenids.
Again, this is significant because hyenas are likely to have left far less flesh on a carcass than a felid such as a sabretooth enabling contemperaneous hominins to benefit \citep{pobiner2015new}.
The use of tools and the cooperative nature of hominins meant they could likely take a certain part of their energetic requirements from carrions depending on their environment \citep{moleon2014humans}. %TG: I've removed the intelligence bit from the sentence.

@@

Further, much work has focused on the existence of scavenging in dinosaurs by using simple energetics approaches that typically focused on a single species namely \textit{T. rex} \citep{ruxton2003could,carbone2011intra} but a recent modelling study investigated the likely prevalence of scavenging across a range of body sizes.
In it the authors demonstrated that species of intermediate body masses (approximatively 500 kg) would have gained the most benefit from scavenging.
This was the result of gut capacity limitations and the effects of competition at the carcass.
At the larger extreme this owes to the fact that gut capacity doesn't scale isometrically with body mass so the benefits of greater mass level off; there's only so much food an individual can consume at a single sitting \citep{calder1996size}.
For the smaller species, larger competitors would have prevented their access to carrion.

In addition to reducing locomotory costs we would expect adaptations that reduce energetic costs of maintenance to be selected for in scavengers because it would maximise the benefit derived from such a sporadic food source. 
Extant reptiles possess an advantage here, in that over the course of a year their food requirements can be 30 times lower than an endotherm of equal size \citep{Nagy1621}.
\cite{devault2002scavenging} suggest this is an avenue for scavenging in snakes because they ``exhibit exceedingly low maintenance metabolisms, and most can survive on  a few scant feedings per year.
It is, therefore, possible for snakes to rely largely on infrequent, less energy-rich meals."
In the same review the authors found occurrences of scavenging spread across five families of snakes and stated that this behaviour is ``far more common than currently acknowledged."\citep{devault2002scavenging}.
The same reasoning can be applied to crocodiles and their allies \citep{forrest2003evidence}. 
A sit and wait strategy is viable for an ectotherm. 
This low existence cost is also realised in many sharks who have coupled low locomotory costs with an ectothermic metabolism. 
The upshot is that 30 kg of blubber can sustain a white shark for over six weeks \citep{carey1982temperature}.

Large body size confers substantial dominance and starvation-resistance benefits \citep{ruxton2004obligate}.
Thus, we would expect scavengers to have this trait selected for even in the case of weight-constrained fliers.
Wandering Albatrosses (\textit{Diomedea exulans}), cinereous vultures (\textit{Aegypius monachus}) and condors (\textit{Vultur gryphus}, \textit{Gymnogyps californianus}) all have body masses that can exceed 10 kg and represent some of the heaviest bird species capable of flight \citep{weimerskirch1992reproductive,ferguson2001raptors,donazar2002effects}.
And as we have noted the Azhdarchid pterosaurs were far bigger again, with estimated body masses of over 200 kg \citep{witton2010size}.
Although \cite{witton2008reappraisal} argued that neck inflexibility and straight, rather than hooked jaw morphology points against pterosaurs existing as \textit{obligate} scavengers, Azhdarchid terrestrial proficiency indicates they would have been comfortable foraging on the ground.
Indeed, extant Marabou Storks (\textit{Leptoptilos crumenifer}) have a comparable morphology and are noted facultative scavengers \citep{monadjem2012survival} so it is reasonable to believe that certain pterosaurs behaved similarly.

@@

@@

%TG: Two points here: one speculation for the non-scavenging behaviour could be because of diet.
Conversely, entire clades appear to lack many, if not all, of these phenotypic adaptations. %TG: I've tried some transition here
For example, the extant bats appear to lack most of the features we have identified as important in reducing handling time.
The larger forms (which are better suited for scavenging, following our previous arguments) are typically frugivores and therefore lack the adaptations for digesting meat.
While the smaller carnivorous bats are mainly found in the microbats which are insectivorous \citep{aguirre2003implications}.
Additionally, their poor terrestrial ability would also count against them when feeding at a carcass \citep{riskin2006terrestrial}.
% Maybe they scavenge on dead insects but here we're talking about scavenging on real meat right?

@@

%TG: maybe change transition here.
Staying in the aquatic setting, the phenomenon of occasional bounties of carrion in the form of whale falls has led some researchers to investigate if a scavenger could survive by seeking out these remains exclusively.
\cite{ruxton2005searching} argued that although this is energetically feasible it's ecologically unlikely.
Any animal that could find such whale carcasses is unlikely to have ignored other types of carrion.
Although no aquatic species have ever exceeded the size of whales, some enormous animals have evolved in this environment before the evolution of whales, including \textit{Leedsichthys}, a bony fish from the Middle Jurassic (174.1-163.5 Mya), that weighed in excess of 20 tonnes.
Thus, the energetic feasibility of a marine scavenger that specialises on large carcasses has a long history.
One point of interest is that of the whaling industry, which provided a bonanza of floating carcasses especially during the 20th century \citep{Whitehead415}.
This meant killer whales could switch from hunting to scavenging, a switch made that much easier by the noise of the whaling vessels that would effectively ring the ``dinner-bells" \citep{Whitehead415}.

>>>>

%KH: need paragraph for how do we know something is (was) a scavenger
\section*{Encounter Rate}
All foraging processes depend on the encounter rate between consumer and resource. 
In the simplest case, this rate can be thought of in terms of a gas diffusion model where the movement of two agents depends only on their relative speed.
As carcasses are stationary, the relative speed between a scavenger and carrion is only dependent on the movement of the scavenger. 
Vertebrates do more than simply bump into resources like gas molecules though, because they can actively detect them through their sensory abilities. 
As such, scavenging potential is strongly affected by search rates, which are determined by both species' physiology and the dimensionality of the environment \citep{pawar2012dimensionality}.

\subsection*{Locomotion}
Scavenging depends more on the ability to efficiently move over large areas than does predation because of the inherent unpredictability of carrion. 
This generally requires an efficient transfer of metabolic energy into movement which relies on both physiology (i.e. metabolism) and the medium of the environment in which the animal is moving (i.e. aerial, aquatic or terrestrial).
Perhaps the most efficient form of locomotion in vertebrates is, paradoxically, found in flying species. 
Despite the energetic costs of flight, the only known vertebrate obligate scavengers are the old and the new world vultures. 
And, although powered flight is energetically expensive, species like vultures can exploit air currents using their large wingspans which allows them to soar at a cost of only twice their metabolic rate \citep{hedenstrom1993migration,spivey2014analysing}. %TG: I didn't get the meaning of this sentence, is soaring doubling the metabolic rate? or is it a word misplacement ("while only doubling their metabolic rate")? and if so, how does that make the point?
By depending on thermal air flows these species can forage over vast ranges \citep{spiegel2013factors}. 
An analagous mode of locomotion is also exploited by seabirds, who use strong ocean winds to search large areas of the oceans \citep{norberg2012vertebrate,thaxter2012seabird}. 
While many species of seabird are likely primarily predators, it seems that albatrosses, who can range many hundreds of kilometres, take a substantial amount of carrion in their diet \citep{croxall1994dead}. 
This is typically in the form of squid, which float on the surface, allowing the birds to readily pluck their remains out of the water \citep{croxall1994dead}.

@@

Similar to aerial species, aquatic scavengers have a locomotory benefit because water is a medium that is conducive to low-cost movement \citep{tucker1975energetic}. 
In fact, the total cost of movement (again in J kg\textsuperscript{-1} m\textsuperscript{-1}) in fish is lower than either running or flying where it scales according to 2.15 $\times$ body mass (kg) \textsuperscript{-0.25} \citep{williams1999evolution} with only \textit{soaring} flight likely to surpass it. 
%KH(I assume the reference is for powered flight, again this needs to be more precise).  
This has led some researchers to argue for the likelihood of an obligate scavenging fish \citep{ruxton2004energetic,ruxton2005searching}. 
Interestingly, style of swimmming in fish does not significantly affect the cost of movement \citep{williams1999evolution}. 
Though sharks perhaps best resemble the large soaring fliers as they depend on large pectoral fins in order to maintain lift as they swim. 
Many shark species show large ranges (e.g. great white sharks (\textit{Carcharodon carcharias})) \citep{bruce2006movements} and it seems reasonable that they would use oceangraphic currents to further reduce movement costs  \citep{ruxton2004energetic}. %and are able to maintain a type of endothermy from thier movements \citep{bernal2015sharks}. %TG: I guess the following is a comment: (I know the Greenland shark is the slowest swimmer). 
Facultative scavenging is seen in many marine groups, including species of extant sharks like white sharks \citep[known to feed on whale carcasses;][]{fallows2013white}, Greenland sharks \citep[feeding on seals;][]{watanabe2012slowest}, and sixgill sharks \citep{anderson2016impact}.
There is evidence too of scavening in extinct species, where shark teeth have been found in the remains of dinosaurs, mosasaurs and Pliocene mysticete whales \citep[5.3 to 3.6 Mya;]{schwimmer1997scavenging,ehret2009caught}.

@@

%% AK:  Need to be careful here because this whole idea of sprawling posture being worse off than erect posture is not that well supported in the literature. 
%% Great Transformations in Vertebrate Evolution By Kenneth P. Dial, Neil Shubin, Elizabeth L. Brainerd talks about it at length 
%% So I think this should stay pitched as an endo/ ectothermy thing rather than posture 
Terrestrial environments are the most energetically costly in which to move, which may be due to the low muscular efficiency of running \citep{tucker1975energetic} as well as the relative inefficiency of gas exchange in mammals cf. birds and fish \citep{williams1999evolution}.
The idea that the evolutionary transition in posture from the sprawling gait of reptiles to the erect posture of mammals conferred a huge advantage to the latter is often mentioned.
The purported advantages include benefits in terms of speed, efficiency, muscle effort and manoeuvrability \citep{sullivan2015posture}. 
Despite being intuitive, as \cite{sullivan2015posture} states, most of the hypotheses in favour of this idea remain to be tested in the context of archosaur evolution. 
Metabolic rate however, unquestionably impacts terrestrial species whereby ectotherms such as many modern reptiles, cannot move for sustained periods \citep{bennett1979endothermy}. 
% Species which do not support an upright position such as modern reptiles limit their ability to move for sustained periods \citep{bennett1979endothermy}. 
This is exacerbated by their sprawling gait which results in the phenomenon known as Carrier's constraint such that the animal can't move and breathe at the same time because the lateral movements impedes its lungs \citep{carrier1987evolution}. 
% These constraints are likely to mean that many ectotermic terrestrial species, which cannot maintain an upright posture would depend on a relatively low level of scavenging (I'm not sure this is true though). 
This would also be true of extinct species. 
It is during the Permian (298.9 - 252.17 Mya) that we have the earliest evidence of vertebrate scavenging where a temnospondyl amphibian fed on the carcass of \textit{Varanops}, a predatory synapsid of the time \citep{reisz2006articulated}. 
%TG: I guess this is a comment too: Snakes are a bit different are more effecent at moving I think - lots of scavanging. 
Both these groups are more likely to depend on low energetic costs in order to compensate for low encounter rates due to their inability to move. 
It is with the evolution of endothermy in the therapsid-mammal lineage \citep{clarke2010temperature} that terrestrial vertebrates would have gained the ability to range more widely, a vital component in seeking out carrion.

Modern endothermic mammals can sustain longer periods of energetically expensive activity \citep{bennett1979endothermy} resulting in larger foraging ranges. 
To quantify this effect with a simple example we can turn to some allometric relationships relating sustainable travelling speed to body mass \citep{ruxton2004obligate}.
If we insert these into a foraging radius model \citep{Enstipp2006Energetics} for a 12 hour foraging day it shows that while a 10 kg reptile can range 6.5 km, an equally sized mammal can range nearly 33 km (See appendix). 
For a foraging scavenger, this ability translates into a greater area searched for food.

@@

%TG: missing some transition here. Maybe something short like "Further back in time" or "Previous to that"...
Although the evolution of bipedal movement was significant in that it freed up the forelimbs for other purposes e.g. climbing, tool-use, wing development etc. it does not differ radically in cost from quadrupedal locomotion \citep[and references therein]{williams1999evolution}. 
For instance, \cite{alexander2004bipedal} shows that, in the case of humans, we are more economical than predicted while walking and less so while running according to predicted costs of terrestrial movement.

The best-known terrestrial bipeds are the dinosaurs and unsurprisingly, given their enduring appeal, the prevalence of scavenging has been explored in the carnivorous theropods.
These were the dominant terrestrial carnivores for most of the Mesozoic Era (252.17 - 66 Mya) and ranged from the chicken-sized to the whale-sized, all of which were bipedal.
They are quite alien to anything we know today which restricts our ability to understand their ecology far more so than extinct mammals \citep{weishampel2004dinosauria}.
Of relevance, are the questions that still persist about their metabolism, with the latest evidence suggesting they were mesothermic i.e. intermediate to ecto- and endotherms \citep{grady2014evidence}. 
We do know that they walked with the erect gait of mammals or birds rather than the sprawling gait of lizards and that they were most likely facultative scavengers \citep{weishampel2004dinosauria,depalma2013physical}. %TG: and cite "Kane et al in press" no? I think these people did some kind of work on that matter.
Taken together, this implies dinosaurs had a foraging range that fell in between that of modern terrestrial mammals and reptiles.

@@

\subsection*{Detection}
It would be pointless to have incredible ranging abilities and not have the sensory architecture to benefit from it.
%If we came at this from a position of complete ignorance we would predict scavengers to have well-developed senses and indeed, this is what we find. %TG: I don't really like this sentence. How about:
As predicted by the necessity of an increased encounter rate, scavengers have well-developed senses.
A simplification of terrestrial, vertebrate scavengers in sensory terms is one of them existing in a two-dimensional plane while foraging for carrion directly.
They can detect carcasses at a range that is defined by the radius of their sensory organs. %, usually the visual and olfactory senses.
As a consequence, they have a much more restricted view of the landscape than do aerial foragers.
Hyenas make up for this in their ability to smell a rotting carcass 4 km away and to hear the vocalisations of conspecifics at a distance of 10 km \citep{mills1989comparative}. 
% Using the approach of \cite{spiegel2013factors} we estimate a spotted hyena %TG: do we need a (\textit{Crocuta crocuta}) here?
%could resolve a 2 metre target at 1 km distance. 
We can compare this to the energetics approach of \cite{ruxton2004obligate}, who calculated a terrestrial scavenger needs to be able to detect carrion at 500 meters in order to survive, which is clearly within the ability of hyenas.  
% While considering prehistoric habitats \cite{ruxton2004obligate} calculated that ``a 1 tonne mammal or reptile, in an ecosystem yielding carrion at densities similar to the current Serengeti, could have met its energy requirements if it could detect carrion over a distance of the order of 400–500 m".
Moreover, the senses of many extant (and in all probability extinct) carnivores meet this required distance, making scavenging feasible for terrestrial species \citep{farlow1994speculations,mech2010wolves}.

@@

Having a panoramic view also means being able to gather a wealth of information from other foragers, be they conspecifics or otherwise \citep{jackson2008effect}.
Again, returning to vultures, the genus \textit{Gyps} consists of highly social and colonially nesting species \citep{fernandez2015density}.
These behaviours allow them forage far more efficiently because one bird can scrounge information on the location of food from another successful forager \citep{KaneVul}.
This efficiency has been exploited by mammals such as hyenas who are known to follow groups of vultures \citep{jones2015african}.

@@

\section*{Handling Time}
Since carrion is not dispatched directly, often the most easily accessible and choicest components of the carcass will be missing or, if present, will be fought over.
Being able to extract nutrients from remnants gives a scavenger a great advantage.
%Thus, the bone crushing ability of hyenas and others reveals another useful scavenger trait. %TG: how about the following for a smoother transition:
Thus, some phenotypic adaptations such as the ability to crush bone (as seen in hyenas) reveal another useful scavenger trait.
Osteophagy is known across a range of terrestrial carnivores and given that some fat-rich mammalian bones have an energy density (6.7 kJ/g) comparable with that of muscle tissue, it makes skeletal remains an enticing resource \citep{brown1989study}.
This ability reached its zenith among hyenas with the evolution of the 110 kg \textit{Pachycrocuta brevirostris} during the Pliocene \citep[3.6 - 2.58 Mya; ][]{palmqvist2011giant}.
Indeed, their extinction has been blamed on the decline of sabretooth cats (Machairodontinae), the unique skull morphology of the latter meant they would leave a large amount of food on a carcass for would-be scavengers \citep{palmqvist2011giant}. 
Earlier in the evolution of mammals, the bone-crushing dogs that evolved during the Oligocene (Borophaginae; 33.9 - 23.03 Mya) have also been compared to hyenas in terms of their feeding ecology \citep{van2003chapter,martin2016pursuit}.

@@

Interestingly, such comparisons have given insight into the feeding ecology of early hominins who, for instance, had the ability to craft tools for breaking open bones \citep{ARCM:ARCM12084}.
The question of where our ancestors placed on the hunter-scavenger axis during the Plio-Pleistocene has been a matter of debate for years \citep{dominguez2002hunting}.
A recent study investigating potential scavenging opportunities for hominins in Kenya found that, even when discounting bone material, there is a substantial amount of scavengeable %TG: is that an adjective?
meat left on predated remains; sufficient to sustain the requirements of an adult male \textit{Homo erectus} \citep{pobiner2015new}.
In some historical hominin-inhabited areas there were a greater number of felids than hyenids.
Again, this is significant because hyenas are likely to have left far less flesh on a carcass than a felid such as a sabretooth enabling contemperaneous hominins to benefit \citep{pobiner2015new}.
The use of tools and the cooperative nature of hominins meant they could likely get a substantial part of their energetic requirements through scavenging depending on their environment \citep{moleon2014humans}. %TG: I've removed the intelligence bit from the sentence.

@@

We know that large body size confers substantial dominance and starvation-resistance benefits \citep{ruxton2004obligate}. 
As such, theropod dinosaurs, who could get up to 15 tonnes, would seem likely candidates for scavenging. 
Much work has focused on the existence of scavenging in dinosaurs by using simple energetics approaches that typically focused on a single species namely \textit{T. rex} \citep{ruxton2003could,carbone2011intra} but a recent modelling study investigated its prevalence across a range of body sizes.

In their work, the authors demonstrated that species of \textit{intermediate} body masses (approximatively 500 kg) would have gained the most benefit from scavenging.
This was the result of gut capacity limitations and the effects of competition at the carcass.
At the larger extreme this owes to the fact that gut capacity doesn't scale isometrically with body mass so the benefits of greater mass level off; there's only so much food an individual can consume at a single sitting \citep{calder1996size}.
For the smaller species, larger competitors would have prevented their access to carrion.

The support of water allows for many aquatic species to reach large sizes thus granting the benefits. 
For example, white sharks, which grow up to 6 metres long, can be sustained by 30 kg of blubber for over six weeks \citep{carey1982temperature}.

Given the advantages of size, we would expect this trait to be selected for even in the case of weight-constrained scavenging fliers.
Wandering Albatrosses (\textit{Diomedea exulans}), cinereous vultures (\textit{Aegypius monachus}) and condors (\textit{Vultur gryphus}, \textit{Gymnogyps californianus}) all have body masses that can exceed 10 kg and represent some of the heaviest bird species capable of flight \citep{weimerskirch1992reproductive,ferguson2001raptors,donazar2002effects}.
And as we have noted the Azhdarchid pterosaurs were far bigger again, with estimated body masses of over 200 kg \citep{witton2010size}.
Although \cite{witton2008reappraisal} argued that neck inflexibility and straight, rather than hooked jaw morphology points against pterosaurs existing as \textit{obligate} scavengers, Azhdarchid terrestrial proficiency indicates they would have been comfortable foraging on the ground.
Indeed, extant Marabou Storks (\textit{Leptoptilos crumenifer}) have a comparable morphology and are noted facultative scavengers \citep{monadjem2012survival} so it is reasonable to believe that certain pterosaurs behaved similarly.

@@

In addition to reducing locomotory costs we would expect adaptations that reduce energetic costs of maintenance to be selected for in scavengers because it would maximise the benefit derived from such a sporadic food source. 
Extant reptiles possess an advantage here, in that over the course of a year their food requirements can be 30 times lower than an endotherm of equal size \citep{Nagy1621}.
\cite{devault2002scavenging} suggest this is an avenue for scavenging in snakes because they ``exhibit exceedingly low maintenance metabolisms, and most can survive on  a few scant feedings per year.
It is, therefore, possible for snakes to rely largely on infrequent, less energy-rich meals."
In the same review the authors found occurrences of scavenging spread across five families of snakes and stated that this behaviour is ``far more common than currently acknowledged."\citep{devault2002scavenging}.
The same reasoning can be applied to crocodiles and their allies \citep{forrest2003evidence} because a sit and wait strategy is viable for an ectotherm. 
%This low existence cost is also realised in many sharks who have coupled low locomotory costs with an ectothermic metabolism. 
%

@@

%TG: Two points here: one speculation for the non-scavenging behaviour could be because of diet.
Conversely, entire clades appear to lack many, if not all, of these phenotypic adaptations. %TG: I've tried some transition here
For example, the extant bats appear to lack most of the features we have identified as important in reducing handling time.
The larger forms (which are better suited for scavenging, following our previous arguments) are typically frugivores and therefore lack the adaptations for digesting meat.
While the smaller carnivorous bats are mainly found in the microbats which are insectivorous \citep{aguirre2003implications}.
Additionally, their poor terrestrial ability and cost of movement on the ground would also count against them when feeding at a carcass \citep{riskin2006terrestrial,voigt2012terrestrial}.
% Maybe they scavenge on dead insects but here we're talking about scavenging on real meat right?

@@

%TG: maybe change transition here.
Staying in the aquatic setting, the phenomenon of occasional bounties of carrion in the form of whale falls has led some researchers to investigate if a scavenger could survive by seeking out these remains exclusively.
\cite{ruxton2005searching} argued that although this is energetically feasible it's ecologically unlikely.
Any animal that could find such whale carcasses is unlikely to have ignored other types of carrion.
Although no aquatic species have ever exceeded the size of whales, some enormous animals have evolved in this environment before the evolution of whales, including \textit{Leedsichthys}, a bony fish from the Middle Jurassic (174.1-163.5 Mya), that weighed in excess of 20 tonnes.
So, despite being unlikely, the energetic feasibility of a marine scavenger that specialises on large carcasses has a long history.
One point of interest is that of the whaling industry, which provided a bonanza of floating carcasses especially during the 20th century \citep{Whitehead415}.
This meant killer whales could switch from hunting to scavenging, a switch made that much easier by the noise of the whaling vessels that would effectively ring the ``dinner-bells" \citep{Whitehead415}.

Perhaps the greatest environmental driver of scavenging tendency is that of temperature. 
The geological record shows the Earth has undergone radical fluctuations in temperature over time.
This will have had a significant bearing on the availability and persistence of carrion.
To illustrate the point, a 10$^{\circ}$C increase in ambient temperature can double carcass decomposition rates \citep{parmenter2009carrion} and geological evidence indicates that the Mesozoic Earth was on average at least 6 $^{\circ}$C warmer than now \citep{sellwood2006mesozoic}.
In terms of specific habitats, it has been shown that decomposition is greater in warm and moist areas versus more xeric ones \citep{beasley2015vertebrates}.
Moreover, oceanic productivity and habitat structure are all impacted by climactic conditions.
The impacts these can have on scavengers have been empirically supported e.g. \cite{beasley2015vertebrates} who point to a series of studies showing how microbes and invertebrates benefit at higher temperatures to the detriment of vertebrate scavengers such that ``above 20$^{\circ}$C vertebrates were able to detect and consume only 19\% of small-mammal carcasses, whereas at temperatures below 18$^{\circ}$C, vertebrates consumed 49\% of carcasses".
This is a sobering thought given the impact we humans are having on the Earth's climate. 
%TG: Really nice to finish on this paragraph + good for transition to conclusion!
